An individual differences perspective on the development of pragmatic reasoning in the preschool years

### 2022-02-25 paper/manuscript.tex
first draft intor for paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{An individual differences perspective on the development of pragmatic reasoning}
\author{Manuel Bohn\textsuperscript{1}, Michael Henry Tessler\textsuperscript{2,3}, Clara Kordt\textsuperscript{4}, Tom Hausmann\textsuperscript{5}, \& Michael C. Frank\textsuperscript{6}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={An individual differences perspective on the development of pragmatic reasoning},
  pdfauthor={Manuel Bohn1, Michael Henry Tessler2,3, Clara Kordt4, Tom Hausmann5, \& Michael C. Frank6},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

@@

\hypertarget{facets-of-pragmatic-reasoning}{%
\section{Facets of pragmatic reasoning}\label{facets-of-pragmatic-reasoning}}

\hypertarget{the-challenges-of-studying-individual-differences}{%
\section{The challenges of studying individual differences}\label{the-challenges-of-studying-individual-differences}}

\hypertarget{the-current-study}{%
\section{The current study}\label{the-current-study}}

@@

The study was presented as an interactive picture book on a tablet computer. The tasks were programmed in \texttt{HTML/JavaScript} and run in a web browser. Pre-recorded sound files were used to address the child (one native German speaker per animal). Children responded by touching objects on the screen. Children were tested in a quiet room in their daycare or in a separate room in a child laboratory. An experimenter guided the child through the study, selecting the different tasks and advancing within each task. In the beginning of the study, children completed a touch training to familiarize themselves with selecting objects. After a short introduction to the different animal characters, children completed the following six tasks. Figure \ref{fig:fig1} shows screenshots for each task and the order in which they were presented.

@@

\caption{Overview of the tasks used in Study 1 to 3. Pictures show screenshots from each task. The vertical order corresponds to the order of presentation in each study. The colors group the tasks along the (Assumed) cognitive processes involved. Blue: utterance-based inferences. Red common ground/discourse-based inferences. Green: Executive functions. Green: Analogical reasoning.}\label{fig:fig1}
\end{figure}

@@

This task was directly taken from Bohn, Le, Peloquin, Köymen, and Frank (2021). Children were told that they were going to visit the animals at their place. The animal greeted the child and told them that they would show them their things. During exposure trials, the child saw three objects from three different categories (e.g., train (vehicle), drum (instrument), orange (fruit); see Figure \ref{fig:fig1}). The animal named of the objects and asked the child to touch it. On the next exposure trial, the child saw three new objects but from the same categories (e.g., bus (vehicle), flute (instrument), apple (fruit)). The animal asked the child to touch the object from the same category as previously (only naming the object, not the category). There were 5 such exposure trials. On the following test trial, the animal used a pronoun to refer to one of the objects (i.e., can you touch \emph{it}). We assumed that children would use the exposure trials to infer that the animal was talking about a certain category and would use this knowledge to identify the referent of the pronoun.Children received five test trials, each with a different category as the target. The position of the objects in exposure trials as well as test trials was pseudo-randomized.

@@

The studies from which we took the mutual exclusivity and informativeness inference tasks also formalized these tasks in an RSA-style model (Bohn et al., 2021, 2022). We refer to this earlier work for more details and a mathematical description of the models. For the present study, we formalized the ad-hoc implicature task within the same RSA framework. All three models shared one common parameter: \(\alpha\). Within the RSA framework, \(\alpha\) has the conceptual role of indicating how informative the listener thinks the speaker is.

This commonality offers a way of relating performance in the three tasks to one another by constraining the three models to use the same value for \(\alpha\). We then used Bayesian inference to estimate the posterior distribution for \(\alpha\) that best explained performance in the three tasks. To adapt this framework to the study of individual differences, we allowed a separate parameter for each participant (\(\alpha_i\)). We estimated \(\alpha_i\) in a hierarchical model as a deviation from a hyper parameter: \(\alpha_i \sim \mathcal{N}(\alpha_j, \sigma^\alpha)\). Given the developmental nature of our data, we defined \(\alpha_j\) via a linear regression as a function of the child's age (\(age_i\)): \(\alpha_j = \beta^\alpha_0 + age_i \cdot \beta^\alpha_1\). Thus, the participant specific value for \(\alpha\) was not only constrained by the performance in the three tasks but also by the child's age.

@@

Our focal model in the confirmatory factor analysis including a latent factor for pragmatic reasoning fit the data well (PPP = 0.50) and with a WAIC of 1,753.45 (se = 32.02, weight = 0.74) better compared to the two alternative models (individual factors model: PPP = 0.51, WAIC = 1,756.48, se = 32.51, weight = 0.16; one factor model: PPP = 0.36, WAIC = 1,758.10, se = 32.37, weight = 0.07). Figure \ref{fig:fig5} shows that factor loadings for the individual tasks as well as their residual variance. The latent pragmatic reasoning factor best explained the mutual exclusivity task, followed by the ad-hoc implicature and the informativeness inference task. The correlation between pragmatic reasoning and executive functions (indicated by the card sorting task) was estimated to be reliably different from zero (\emph{r} = 0.33; model estimate = 0.20, 95\% CrI {[}0.02 - 0.39{]}). There was no systematic relation between pragmatic reasoning and analogical reasoning (as indicated by the relational match-to-sample task): \emph{r} = 0.06; model estimate = 0.04, 95\% CrI {[}-0.11 - 0.18{]}. However, the latter result should be taken with a groin of salt given the unknown psychometric properties of the relational match-to-sample task.

The results of the analysis based on the cognitive model yield similar conclusions compared to the confirmatory factor analysis. Participant specific speaker informativeness parameters (\(\alpha\)) were correlated with performance in the card sorting (\emph{r} = 0.31, 95\% CI{[}0.15 - 0.47{]}) but not the relational match-to-sample task (\emph{r} = 0.03, 95\% CI{[}-0.15 - 0.20{]})

@@

Present six, reliable (even when corrected for age) and ready to use tasks that cover a range of phenomena subsumed under pragmatic inference. WE divided them up into utternace based and interaction based and more systematic realtions between utternace based. Nevertehless, the others need to be explored more

Across studies we saw relations between utterance based inference tasks. From different theoretical perspectives, they might be more or less obvious, especially for ME. From our perspective they make a lot of sense - we explicate this view in our model.

Why is there this relation with card sorting? The advantage of using a cognitive model to study individual differences is that the parameters have a more psychologically plausible interpretation compared to data analytic methods. If e take the model for real, we can speculate why card sorting matters. Switching rules requires seeing stimuli in different light - focus on different aspect of them, consider alternative ways of seeing them, not go with what you did previously. The inference in the model is driven by not going just with what is true but considering alternative utterances - some commonalities here that might be responsible - need to be modeled explicitly and contrasted to alternative views. (check other theoretical accounts on that)

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-bohn_le_peloquin_koymen_frank_2020}{}}%
Bohn, M., Le, K. N., Peloquin, B., Köymen, B., \& Frank, M. C. (2021). Children's interpretation of ambiguous pronouns based on prior discourse. \emph{Developmental Science}, \emph{24}(3), e13049.

@@

@@

@@

@@

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{An individual differences perspective on the development of pragmatic reasoning in the preschool years}
\author{Manuel Bohn\textsuperscript{1}, Michael Henry Tessler\textsuperscript{2,3}, Clara Kordt\textsuperscript{4}, Tom Hausmann\textsuperscript{5}, \& Michael C. Frank\textsuperscript{6}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={An individual differences perspective on the development of pragmatic reasoning in the preschool years},
  pdfauthor={Manuel Bohn1, Michael Henry Tessler2,3, Clara Kordt4, Tom Hausmann5, \& Michael C. Frank6},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

@@

Communication predates language. Before children produce their first words, they communicate with the world around them using vocalizations and gestures (Bates, Benigni, Bretherton, Camaioni, \& Volterra, 1979; Bruner, 1974). The process of language learning recruits many of the social-cognitive processes that underlay pre-verbal communication (Bohn \& Frank, 2019; E. V. Clark, 2009; Tomasello, 2009). Even for proficient language users is communication not reducible to the words being exchanged. The common thread running through these different aspects of human communication is its inferential nature: what a speaker means -- verbal or otherwise -- is underdetermined by the parts that make up the utterance. It takes contextual social inferences to recover the intended meaning (H. H. Clark, 1996; Grice, 1991; Levinson, 2000; Sperber \& Wilson, 2001). These contextual social inferences are often referred to as \emph{pragmatic} inferences (e.g., Bohn \& Frank, 2019).

The development of pragmatics has been widely studied in recent years (for a recent review see Bohn \& Frank, 2019). This research covers a range of different phenomena ranging from so-called \emph{pure pragmatics} (Matthews, 2014) in non-verbal communication in infancy to sophisticated linguistic inferences developing much later (Huang \& Snedeker, 2009; Papafragou \& Skordos, 2016). A growing portion of this work is devoted to studying individual differences (Matthews, Biney, \& Abbot-Smith, 2018; E. Wilson \& Katsos, 2021). The motivation behind this is twofold: on the one hand, individual differences offer insights into the underlying psychological processes (Kidd, Donnelly, \& Christiansen, 2018; Matthews et al., 2018; A. C. Wilson \& Bishop, 2022). On the other hand, deficits in pragmatic abilities have been linked to behavioral problems and forms of language impairment (Helland, Lundervold, Heimann, \& Posserud, 2014).

In their recent review of individual differences in children's pragmatic ability, Matthews et al. (2018) identified three issues that significantly limit what we can learn from individual differences research. First, most studies have insufficient sample sizes so that small and medium sized correlations among pragmatics tasks and between pragmatics tasks and measures for other cognitive abilities cannot be reliably detected. Second, the tasks used to assess pragmatic abilities often have poor or unknown psychometric properties. For example, many tasks only have a single trial and are therefore unable to capture variation between children. Reliability is not assessed, making it unclear if the task captures stable characteristics (see also Russell \& Grizzle, 2008). Third, the cognitive processes underlying pragmatic inferences in a particular task are underspecified. As a consequence, there is no clear rationale for why a task should correlate with another cognitive measure.

The studies presented here directly address the issues identified by Matthews and colleagues (Matthews et al., 2018). We identified six pragmatic reasoning tasks in children between three and five years of age and investigated their psychometric properties, in particular their re-test reliability. Reliable tasks are a necessary precondition for meaningful individual differences research (Fried \& Flake, 2018; Hedge, Powell, \& Sumner, 2018). Next, we investigated the relations among different pragmatic reasoning tasks as well as between pragmatic reasoning and executive functions in a sample large enough to detect small to medium sized correlations. Finally, we introduced computational cognitive models of pragmatic reasoning to the study of individual differences. This allowed us to formalize the cognitive processes we think underlay pragmatic reasoning and thereby provide a substantive theoretical account for why certain pragmatic reasoning tasks are related to one another and why executive functions relate to them.

The six tasks we selected were developmental adaptations of referential communication games inspired by research in experimental pragmatics (Noveck \& Reboul, 2008; Noveck \& Sperber, 2004). They all share a common trial-by-trial structure in which the test event always involved an agent producing an ambiguous utterance that children had to resolve using pragmatic reasoning. This allowed us to run multiple trials per task, and with that, to capture individual differences. We grouped the tasks into two broad categories (Figure \ref{fig:fig1}). On the one hand, utterance-based tasks asked children to derive inferences from the words and gestures the speaker produced in context. On the other hand, common ground/discourse-based tasks asked children to derive inferences from the social interaction that preceded the utterance.

For the utterance-based category, we selected a mutual exclusivity, informativeness inference, and ad-hoc implicature task. Mutual exclusivity -- or principle of contrast -- describes the phenomenon that children tend to map a novel word to an unknown object (Bion, Borovsky, \& Fernald, 2013; E. V. Clark, 1988; Halberda, 2003; Lewis, Cristiano, Lake, Kwan, \& Frank, 2020; Markman \& Wachtel, 1988; Merriman, Bowman, \& MacWhinney, 1989). Informativeness inferences describe situations in which children identify the referent of a novel word by assuming that the speaker is trying to be informative. Being informative translates to using words that have smaller extensions in context (Frank \& Goodman, 2014). Ad-hoc implicature describes inferences that ask the child to contrast an utterance with alternatives that the speaker could have used but did not (Katsos \& Bishop, 2011; Stiller, Goodman, \& Frank, 2015; Yoon \& Frank, 2019).

For the discourse-based category, we selected a speaker preference, discourse novelty and discourse continuity task. In the speaker preference task, the child had to track the preference of a speaker in order to identify the referent of a novel word (Saylor, Sabbagh, Fortuna, \& Troseth, 2009). Discourse novelty refers to a situation in which the child tracks the temporal appearance of objects in context and expects the speaker to refer to objects that are new in context (Akhtar, Carpenter, \& Tomasello, 1996; Diesendruck, Markson, Akhtar, \& Reudor, 2004). In the discourse continuity task, the child had to infer and track the topic of an ongoing conversation to resolve ambiguity (Akhtar, 2002; Bohn, Le, Peloquin, Köymen, \& Frank, 2021).

In addition to the pragmatics tasks, we also included a tasks for executive functions (Zelazo, 2006) and of analogical reasoning (Christie \& Gentner, 2014). Executive functions refer to a family of top-down mental processes that enable us to inhibit automatic or intuitive responses and allow us to concentrate and focus attention on particulars (Diamond, 2013). A substantial body of research has investigated the link between executive functions and pragmatics -- with mixed results (Matthews et al., 2018; Nilsen \& Graham, 2009). Analogical reasoning refers to the ability to reason about abstract relations between stimuli (Carstensen \& Frank, 2021) -- an ability that has not been linked to pragmatics, at least not to the same extent as executive functions.

After finding in Study 1 and 2 that most of the pragmatics tasks showed good to excellent re-test reliability, we studied a larger sample of children in Study 3 to investigate relations between the three utterance-based tasks. We focused on these tasks for theoretical reasons: we assume that -- computationally -- they share a common contrastive inference process according to which the listener compares the utterance the speaker made to alternative possible utterances. The listener expects the speaker to be helpful and informative, that is, to choose the utterance that best communicates the intended message. We formalize these assumptions in a computational cognitive model which we then use to study individual differences in this alleged process. Study 3 also included tasks for executive functions and analogical reasoning. Across analytical approaches, we found systematic relations among the pragmatics tasks as well as between pragmatics and executive functions, but not analogical reasoning.

@@

\caption{Overview of the tasks used in Study 1 to 3. Pictures show screenshots from each task. The vertical order corresponds to the order of presentation in each study. The colors group the tasks along the (Assumed) cognitive processes involved. Blue: utterance-based inferences. Red common ground/discourse-based inferences. Green: Executive functions. Yellow: Analogical reasoning.}\label{fig:fig1}
\end{figure}

The study was presented as an interactive picture book on a tablet computer (Frank, Sugarman, Horowitz, Lewis, \& Yurovsky, 2016). The tasks were programmed in \texttt{HTML/JavaScript} and run in a web browser. Pre-recorded sound files were used to address the child (one native German speaker per animal). Children responded by touching objects on the screen. Children were tested in a quiet room in their daycare or in a separate room in a child laboratory. An experimenter guided the child through the study, selecting the different tasks and advancing within each task. In the beginning of the study, children completed a touch training to familiarize themselves with selecting objects. After a short introduction to the different animal characters, children completed the following six tasks. Figure \ref{fig:fig1} shows screenshots for each task and the order in which they were presented.

@@

@@

This task was directly taken from Bohn et al. (2021). Children were told that they were going to visit the animals at their place. The animal greeted the child and told them that they would show them their things. During exposure trials, the child saw three objects from three different categories (e.g., train (vehicle), drum (instrument), orange (fruit); see Figure \ref{fig:fig1}). The animal named of the objects and asked the child to touch it. On the next exposure trial, the child saw three new objects but from the same categories (e.g., bus (vehicle), flute (instrument), apple (fruit)). The animal asked the child to touch the object from the same category as previously (only naming the object, not the category). There were 5 such exposure trials. On the following test trial, the animal used a pronoun to refer to one of the objects (i.e., can you touch \emph{it}). We assumed that children would use the exposure trials to infer that the animal was talking about a certain category and would use this knowledge to identify the referent of the pronoun.Children received five test trials, each with a different category as the target. The position of the objects in exposure trials as well as test trials was pseudo-randomized.

@@

The studies from which we took the mutual exclusivity and informativeness inference tasks also formalized these tasks in an RSA-style model (Bohn et al., 2021, 2022). We refer to this earlier work for more details and a mathematical description of the models. For the present study, we formalized the ad-hoc implicature task within the same RSA framework. The three models share the same contrastive inference process according to which the listener compares the speaker's utterance to alternative, possible ones. The speaker then expects the listener to be informative (with degree \(\alpha\)), that is, choose the utterance that best communicates the intended message. In the mutual exclusivity task, the alternative utterance would be to use a familiar word. In the case of the informative inference task, the alternative was point to the object with only one property. For the ad-hoc implicature task it was to single out the property that was shared by both objects. In all cases, these alternative utterances would be better suited to communicate about the respective other referent.

As noted above, all models shared one common parameter: the speaker informativeness parameter \(\alpha\). This commonality offers a way of relating performance in the three tasks to one another by constraining the three models to use the same value for \(\alpha\). We then used Bayesian inference to estimate the posterior distribution for \(\alpha\) that best explained performance in the three tasks. To adapt this framework to the study of individual differences, we allowed a separate parameter for each participant (\(\alpha_i\)). We estimated \(\alpha_i\) in a hierarchical model as a deviation from a hyper parameter: \(\alpha_i \sim \mathcal{N}(\alpha_j, \sigma^\alpha)\). Given the developmental nature of our data, we defined \(\alpha_j\) via a linear regression as a function of the child's age (\(age_i\)): \(\alpha_j = \beta^\alpha_0 + age_i \cdot \beta^\alpha_1\). Thus, the participant specific value for \(\alpha\) was not only constrained by the performance in the three tasks but also by the child's age.

@@

Next, we turn to the results of the confirmatory factor analysis. Our focal model -- including a latent factor for pragmatic reasoning -- fit the data well (PPP = 0.50) and with a WAIC of 1,753.45 (se = 32.02, weight = 0.74) better compared to the two alternative models (individual factors model: PPP = 0.51, WAIC = 1,756.48, se = 32.51, weight = 0.16; one factor model: PPP = 0.36, WAIC = 1,758.10, se = 32.37, weight = 0.07). Figure \ref{fig:fig5} shows that factor loadings for the individual tasks as well as their residual variance. The latent pragmatic reasoning factor best explained the mutual exclusivity task, followed by the ad-hoc implicature and the informativeness inference task. The correlation between pragmatic reasoning and executive functions (indicated by the card sorting task) was estimated to be reliably different from zero (\emph{r} = 0.33; model estimate = 0.20, 95\% CrI {[}0.02 - 0.39{]}). There was no systematic relation between pragmatic reasoning and analogical reasoning (as indicated by the relational match-to-sample task): \emph{r} = 0.06; model estimate = 0.04, 95\% CrI {[}-0.11 - 0.18{]}. However, the latter result should be taken with a grain of salt given the unknown psychometric properties of the relational match-to-sample task.

Finally, we present the results of the cognitive modeling analysis. Using the data from Study 2, we could show that participant specific speaker informativeness parameters (\(\alpha\)) were highly reliable (Figure \ref{fig:fig6}B). The scale parameter suggested that the mutual exclusivity task was easier and the informativeness inference task was harder compared to the ad-hoc implicature task (Figure \ref{fig:fig6}C). When correlating \(\alpha\) with performance in the other two tasks, the cognitive modeling approach yielded similar conclusions compared to the confirmatory factor analysis (Figure \ref{fig:fig6}A): There was a substantial correlation with the card sorting (\emph{r} = 0.31, 95\% CI{[}0.15 - 0.47{]}) but not the relational match-to-sample task (\emph{r} = 0.03, 95\% CI{[}-0.15 - 0.20{]}). The same limitations apply to the latter result as for the confirmatory factor analysis.

@@

In this paper, we explored the development of pragmatic inferences in the preschool years. We identified six tasks covering a broad range of pragmatic phenomena. We found them to have good psychometric properties, in particular good to excellent re-test reliability. We then selected three utterance-based inference tasks for well-powered study of relations among different types of pragmatic abilities and between pragmatics and other cognitive processes. The results showed systematic relations between the tasks. We used a computational cognitive model of pragmatic reasoning to formalize the cognitive processes we believed the tasks to share. Finally, we found pragmatic abilities to be related to a task of executive functions.

One of the main contributions of this paper is that if presents six pragmatic inference tasks that are highly robust and reliable. Whenever we used a task in two studies (mutual exclusivity, informativeness inference, ad-hoc implicature), we found the developmental results to closely replicate. in Study 1 and 2, all tasks showed good re-test reliability -- even when corrected for age. A notable exception was the informativeness inference task in Study 1. However, after making some procedural changes, it turned out to be robust and reliable as well. Taken together, the tasks are highly suitable for individual differences research. The material is freely available via the online repository associated with the study.

We grouped the tasks into utterance-based and common ground/discourse based. This grouping broadly captured the kind of information that we assumed to be relevant to compute the inference. For Study 3, we focused on the three utterance-based tasks. The main reason was theoretical. We were able to build on earlier work and formalize the inferences involved in these tasks in a common computational framework. We specified the structural overlap between the tasks and identified a parameter in the model that we used to capture individual differences. The shared structural features involve a recursive social inference process according to which the listener expects the speaker to select the most informative of a set of possible utterances. The individual difference parameter captured how informative the listener expected the speaker to be\footnote{this interpretation follows from the conceptual role the parameter plays in the RSA framework. Technically, it acts as an amplifier of the basic inference}. Previous accounts, in particular of mutual exclusivity (reviewed in Lewis et al., 2020), would not have predicted such an overlap. The results of Study 3 (and also Study 2) supported our assumptions as we found systematic relations between the three tasks.

Our formal model also allows us to speculate about why we saw a systematic relation across the three studies between pragmatic inference and the card sorting task as a measure of executive functions. Before we do so, we want to emphasize that the model is first and foremost a computational description of the tasks an not a model of a psychological process (cf. Goodman \& Frank, 2016). For the following, we assume more psychological realism than perhaps warranted. The card sorting task asks the child to switch between rules half way. This requires inhibiting a previously dominant response and attending to different features of the cards. The pragmatic inference in the model involves contrasting the observed utterance with alternative utterances. This could be described as requiring inhibiting interpretations that follow from other plausible interpretations and considering that there would have been better ways of saying it?? However should find a common representation and model this explicitly.

Switching rules requires seeing stimuli in different light - focus on different aspect of them, consider alternative ways of seeing them, not go with what you did previously. The inference in the model is driven by not going just with what is true but considering alternative utterances - some commonalities here that might be responsible - need to be modeled explicitly and contrasted to alternative views. (check other theoretical accounts on that).

We did not look at the other tasks in the same level of detail

Cartoon picture book presentation - not real speakers. But results match those of earlier in personn studies and generally tablet based tasks seem to be fine.

Studies only weird kids from one cultural background, Nevertheless the tasks had been developed in eglish and were now used in German - suggets some portability. More needs to be done obviously.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-akhtar2002relevance}{}}%
Akhtar, N. (2002). Relevance and early word learning. \emph{Journal of Child Language}, \emph{29}(3), 677--686.

\leavevmode\vadjust pre{\hypertarget{ref-akhtar1996role}{}}%
Akhtar, N., Carpenter, M., \& Tomasello, M. (1996). The role of discourse novelty in early word learning. \emph{Child Development}, \emph{67}(2), 635--645.

\leavevmode\vadjust pre{\hypertarget{ref-bates1979emergence}{}}%
Bates, E., Benigni, L., Bretherton, I., Camaioni, L., \& Volterra, V. (1979). \emph{The emergence of symbols: Cognition and communication in infancy}. New York: Academic Press.

\leavevmode\vadjust pre{\hypertarget{ref-bion2013fast}{}}%
Bion, R. A., Borovsky, A., \& Fernald, A. (2013). Fast mapping, slow learning: Disambiguation of novel word--object mappings in relation to vocabulary learning at 18, 24, and 30 months. \emph{Cognition}, \emph{126}(1), 39--53.

\leavevmode\vadjust pre{\hypertarget{ref-bohn2019pervasive}{}}%
Bohn, M., \& Frank, M. C. (2019). The pervasive role of pragmatics in early language. \emph{Annual Review of Developmental Psychology}, \emph{1}(1), 223--249.

\leavevmode\vadjust pre{\hypertarget{ref-bohn_le_peloquin_koymen_frank_2020}{}}%
Bohn, M., Le, K. N., Peloquin, B., Köymen, B., \& Frank, M. C. (2021). Children's interpretation of ambiguous pronouns based on prior discourse. \emph{Developmental Science}, \emph{24}(3), e13049.

@@

\leavevmode\vadjust pre{\hypertarget{ref-bruner1974communication}{}}%
Bruner, J. S. (1974). From communication to language---a psychological perspective. \emph{Cognition}, \emph{3}(3), 255--287.

@@

\leavevmode\vadjust pre{\hypertarget{ref-carstensen2021graded}{}}%
Carstensen, A., \& Frank, M. C. (2021). Do graded representations support abstract thought? \emph{Current Opinion in Behavioral Sciences}, \emph{37}, 90--97.

@@

\leavevmode\vadjust pre{\hypertarget{ref-clark1988logic}{}}%
Clark, E. V. (1988). On the logic of contrast. \emph{Journal of Child Language}, \emph{15}(2), 317--335.

\leavevmode\vadjust pre{\hypertarget{ref-clark2009first}{}}%
Clark, E. V. (2009). \emph{First language acquisition}. Cambridge: Cambridge University Press.

\leavevmode\vadjust pre{\hypertarget{ref-clark1996using}{}}%
Clark, H. H. (1996). \emph{Using language}. Cambridge: Cambridge University Press.

\leavevmode\vadjust pre{\hypertarget{ref-diamond2013executive}{}}%
Diamond, A. (2013). Executive functions. \emph{Annual Review of Psychology}, \emph{64}, 135--168.

\leavevmode\vadjust pre{\hypertarget{ref-diesendruck2004two}{}}%
Diesendruck, G., Markson, L., Akhtar, N., \& Reudor, A. (2004). Two-year-olds' sensitivity to speakers' intent: An alternative account of samuelson and smith. \emph{Developmental Science}, \emph{7}(1), 33--41.

@@

\leavevmode\vadjust pre{\hypertarget{ref-frank2016using}{}}%
Frank, M. C., Sugarman, E., Horowitz, A. C., Lewis, M. L., \& Yurovsky, D. (2016). Using tablets to collect data from young children. \emph{Journal of Cognition and Development}, \emph{17}(1), 1--17.

\leavevmode\vadjust pre{\hypertarget{ref-fried2018measurement}{}}%
Fried, E. I., \& Flake, J. K. (2018). Measurement matters. \emph{APS Observer}, \emph{31}(3).

@@

\leavevmode\vadjust pre{\hypertarget{ref-grice1991studies}{}}%
Grice, H. P. (1991). \emph{Studies in the way of words}. Cambridge, MA: Harvard University Press.

\leavevmode\vadjust pre{\hypertarget{ref-halberda2003development}{}}%
Halberda, J. (2003). The development of a word-learning strategy. \emph{Cognition}, \emph{87}(1), B23--B34.
